Automated verification of electrical systems: literature review and case studies of stand-alone solar photovoltaics\thanks{Supported by Newton Fund (ref. 261881580) and FAPEAM (Amazonas State Foundation for Research Support, calls 009/2017 and PROTI Pesquisa 2018).}

### 2019-01-19 Tex/CAV2019/Trindade_Cordeiro_CAV2019.tex
added comments
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>
paper update

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{makecell}
\usepackage{algorithm,algorithmic}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

@@

\keywords{Automated verification  \and Model Checking \and Electrical systems \and Solar photovoltaic systems.}
\end{abstract}
%
%
%
\section{Introduction}
%%%%%In 2015, 193 Member States of the United Nations agreed as part of the Sustainable Development Goals (SDG) with a specific target to 'ensure access to affordable, reliable and modern energy for all by 2030' – universal access to electricity and clean cooking~\cite{IEAeao2017}. Therefore electricity has a important role, based on two main issues: %that are on the agenda of all current rulers and society in general, regardless of whether it is a developed, emerging or developing country: 
%%%%%the impact on the global climate change and poverty alleviation.
%
%%%%%Related to climate change, the 23rd session of the Conference of the Parties (COP23) to the United Nations Framework Convention on Climate Change (UNFCCC) in 2017 was the second session after the Paris Agreement, which aims to keep the maximum global average temperature rise as close as possible to 1.5 $^{o}$C. %During COP23, with respect to the future market development of renewable power and solar PV (photovoltaics) in particular, it was created an alliance of nations and states committed to moving the world from burning coal to cleaner power sources was a promising sign at COP23 \cite{Jager-Waldau}. 
%%%%%Decarbonization of our energy supply is an important component to achieve the targets, because 65.3\% of the worlds current CO\textsubscript{2} emissions (from energy supple) are due to burning fossil fuels according with 2016 data: 81.1\% of our total primary energy supply depended on burning fossil fuels, namely 27.1\% coal, 31.9\% oil and 22.1\% natural gas~\cite{IEAwes2018}. In terms of final energy consumption electricity only accounted for 18.6\%, but was responsible for 32\% of the total CO\textsubscript{2} emissions~\cite{IEAwes2018}. Photovoltaics (PV) is one of the key technology options for implementing the shift to a decarbonized energy supply, because it is abundant, cannot be monopolized  and can be deployed in a modular way almost everywhere on Earth.
%
%Related to poverty alleviation, 
%
%In reality, the biggest barrier to universal energy access is not the capacity to generate electricity, but rather the ability to get energy to those who need it most. The concept of energy poverty and access tends to focus on very small, incremental shifts in the delivery of energy to poor people. The total energy demand of these shifts constitutes only a small fraction of the ‘gap’ between electricity supply in much of Africa versus more industrialised countries, or the gap between the region’s current and projected electricity supply. https://policy-practice.oxfamamerica.org/static/media/files/FINAL_speakingpowertotruth_SH.pdf
%
%%%%%Poverty, in other hand, needs to be addressed through socio-economic development. Poverty %is conceptualized in material terms as not having access to adequate levels of food, water, clothing, shelter, sanitation, health care and education. This 
%%%%%can be translated into people having insufficient income. 
%
For numerous people around the world, a better life means getting access to basic needs such as food, health services, housing and clean water. None of these basic needs can be provided without energy. Energy is one of the most essential inputs into sustaining people‘s livelihoods. Lack of access to clean and affordable energy is considered a core dimension of poverty~\cite{Hussein2012}.

Progress has been made worldwide. The number of people without electricity access fell below 1 billion for the first time in 2017~\cite{IEAweo2018}. To provide universal electricity for all, decentralized systems, led by solar PV in off-grid and mini-grid systems, will be the least-cost solution for three-quarters of the additional connections needed, but that grid extension will still have a role to play, especially in urban areas~\cite{IEAweo2018}.

@@

In each iteration, Automizer pick a trace (sequence of statements) that leads from the initial location to the error location and check whether the trace is feasible or infeasible. If the trace is feasible, an error is reported to the user; otherwise it is computed a sequence of predicates along the trace as a proof of the trace’s infeasibility. The predicates are a sequence of interpolants since each predicate “interpolates” between the set of reachable states and the set of states from which we cannot reach the error. In the refinement step of the CEGAR loop, Automizer tries to find all traces whose infeasibility can be shown with the given predicates and subtract these traces from the set of (potentially spurious) error traces that have not yet been analyzed. It is used automata to represent sets of traces. The difference to a classical CEGAR-based predicate abstraction is that Automizer have to do any logical reasoning (e.g., SMT solver calls) that involves predicates of different CEGAR iterations~\cite{Automizer2018}. 
%
%----------------------------------------
\section{Automated verification of Stand-alone Solar PV Systems }
\label{sec:Methodology}
%----------------------------------------
\subsection{Stand-alone Solar PV Systems}
PV systems are classified into distinct types~\cite{Mohanty}. Specifically for remote rural areas of developing countries or places where the grid extension is not feasible, the most suitable configuration is the regulated stand-alone system with battery and AC load; this configuration is the focus of case-studies of this work.

The mathematical modeling of the PV system is based on modular blocks, as illustrated in Fig.\ref{fig:blockdiagram}. It identifies the PV generator, batteries, charge controller, inverter, and AC load. The PV generator, which can be a panel or an array, is a semiconductor device that can convert solar energy into DC electricity. In Fig.\ref{fig:blockdiagram}, there are two variables that depend on the site where the system is deployed and the weather (i.e., solar irradiance $G$ and temperature $T$). For night hours or rainy days, we need to hold batteries, where power can be stored and used. The use of battery as a storage form implies the presence of a charge controller~\cite{Hansen}. The PV arrays produce DC and therefore when the PV system contains an AC load, a DC/AC conversion is required. That converter is called of inverter; and the AC load dictates the behavior of AC electrical load from the house that will be fed by the system.
\begin{figure}[h]
\includegraphics[width=0.6\textwidth]{blockdiagramPVS2}
\centering
\caption{Block diagram for a typical stand-alone PV system~\cite{Hansen}.}
\label{fig:blockdiagram} 
\end{figure}
A wide variety of models exists for estimation of power output of PV modules (and $I-V$ or $P-V$ curves). However, the present work will rely on the simplified model of 1-diode. It was demonstrated that it has a small error rate, between 0.03\% and 4.68\% from selected PV panels tested~\cite{Saloux}. The model uses formulas from different references~\cite{Hansen}, \cite{Saloux}, and \cite{Ross}.

To model the batteries, it was used the most common based on lead-acid~\cite{Copetti}. Related to the charge controller, it was used the 4-steps controller processes with MPPT (maximum power point tracking) to decide if the PV array must be disconnected from the system, or if the PV can be reconnected, or if the load is disconnected and when the load is reconnected~\cite{Hansen}. The inverter model is the simplest and is based on efficiency, i.e., the relation between the AC output power and the DC input power of the component~\cite{Hansen}. The AC load, from every case study, is defined by a 24 hour curve who reflects the estimated use of all electrical loads from the house. Load information was collected during a presential survey performed in each house.

@@

Algorithm~\ref{alg:verification-algorithm} describes the pseudo-code used to perform the automated verification. %Line 1 indicates a function call that performs the size checking of the each component of the PV system. %: using Equations \eqref{eq:NTPmin}, \eqref{eq:NTP}, \eqref{eq:NPSmin}, \eqref{eq:NPS}, \eqref{eq:NPP}, and \eqref{eq:NPPmin} to verify the PV panel; using \eqref{eq:Cbank}, \eqref{eq:Nbtotal}, and \eqref{eq:batcheck} to verify the batteries; using \eqref{eq:vcvsystem}, \eqref{eq:icmin}, and \eqref{eq:icicmin} to verify the charge controller; and using \eqref{eq:vindc}, \eqref{eq:voutac}, and \eqref{eq:invcheck} to verify the inverter. 
%The verification is carried out by the \textit{assert} macro from the ANSI-C programming language to encode each equation of sizing check. The argument to the \textit{assert} statement must be \textit{true} if the system specification is met; otherwise, the program aborts and prints a counterexample indicating a property violation. If there is no property violation, then the verification algorithm continues and 
The batteries are assumed to have initial SOC of 80\% (Line 1). %%100% before, i.e., the batteries are charged. That is important because the system is said ready to use when the batteries are charged.
%At both case studies, it was considered 21 years of historic data from Manaus. 
%According with X and Y, t
Information related to average temperature ($T$) and solar irradiance ($G$), maximum and minimum annual, are given to the algorithm in Lines 7 to 10 using non-deterministic variables from ESBMC to explore all possible states and the \textit{assume} macro to constrain the non-deterministic values using a given range. 
%and irradiance varies from 0 W/m$^{2}$ to 852 W/m$^{2}$ (with minimum of 274 W/m$^{2}$ during the daytime, when there is sunlight). 
%there is PV generation only between 8:00 h and 16:00 h every day, 
%with zero electric energy generation from 18:00 h to 6:00 h of the next day; and with insignificant generation from 6:00 h to 8:00 h, and from 16:00 h to 18:00 h of the same day. 
In order to reduce the computational effort of the algorithm,
% caused by the state explosion inherent of the technique, 
every 24h-day was considered as a time-step of 1 hour, and it was split into two parts: (a) one where it is possible to occur PV generation, during daylight, with a duration in hours depending on each site (but dependent on the sun and weather conditions); and (b) one that includes all the remaining day (without any PV generation). Therefore, our approach depends on specific data about the solar irradiation levels to define the average amount of hours of PV generation.

@@

Next, representing the time of the day where PV generation is not possible anymore, starting in Line 22, the algorithm will only discharge the batteries (using the 1 hour step) until a new charging process (at the next day) starts. Specific \textit{assert} in Line 26 will check the state of charging from batteries, and they will violate the property if their levels reach the minimum that represents a discharged battery; therefore, the PV system is unable to supply energy to the house. Nevertheless, if the verification engine does not fail, then we can conclude that the PV system does not need further corrections up to the given bounds.
%
%\textcolor{red}{this sentence is unclear... After this process is started the battery autonomy verification, from line 31}. \textcolor{red}{this sentence is unclear... Based on the fact that won't be PV generation after a given time of the day, the algorithm will only discharge the batteries until a new charging process (at the next day) to start.} \textcolor{red}{what do you mean by The formal verification is guaranteed?...  The formal verification is guaranteed by  macro to specific variables of the model, according lines 27 and 35.}
% and the non-deterministic variables $G$ and $T$ are considered during the formal verification of the system, otherwise, during the other two periods, there is no PV generation and just the power consumption from the backup batteries. 
%Within this 8h-period, $G$ and $T$ are automated verified with different values every one hour.
%, and change their value every 1 h according with the algorithm created using the technique.
 \begin{algorithm}
 \caption{Model checking algorithm for stand-alone PV}
 \begin{algorithmic}[1]
 \renewcommand{\algorithmicrequire}{\textbf{Input:}}
 \renewcommand{\algorithmicensure}{\textbf{Output:}}
  \STATE $SOC \leftarrow 80\%$ \\
%  \COMMENT {Starting with the PV generation time}
% \\ 
 \textit{LOOP Process}
  \FOR {$h = 1$ to $Hours \, of \, PV \, generation$}
  \STATE $G \leftarrow nondet \textunderscore uint(\,)$ \COMMENT {$G$ is non-deterministic variable}
  \STATE $T \leftarrow nondet \textunderscore uint(\,)$ \COMMENT {$T$ is non-deterministic variable}
  \STATE assume ($Gmin \leq G \leq Gmax$) \COMMENT {restricting $G$ values}
  \STATE assume ($Tmin \leq T \leq Tmax$) \COMMENT {restricting $T$ values}
  \STATE $Imax, Vmax \leftarrow PVgenerationMODEL (G,T)$ 
  \\
%  \COMMENT {If-then-else sequence to imitate charge controller work}
  \IF {($battery \, is \, empty$) AND ($PV \, is \, generating$)}
    \STATE charge battery in 1h
    \STATE PV feed the house
  \ELSIF {($battery \, is \, empty$) AND NOT($PV \, is \, generating$)}
    \STATE FAIL with assert macro
  \ELSIF {NOT($battery \, is \, empty$) AND ($PV \, is \, generating$)}
    \STATE stop battery charge
    \STATE PV feed the house
  \ELSIF {NOT($battery \, is \, empty$) AND NOT($PV \, is \, generating$)}
    \STATE discharge battery in 1h
    \STATE Battery feed the house
  \ENDIF
  \STATE $h \leftarrow (h+1)$
  \ENDFOR
 \\ \textit{Start of battery autonomy verification:}
%%  initialization\;
\STATE $AutonomyCount \leftarrow 1$
 \WHILE {$AutonomyCount \leq autonomy$}
  %%\FOR {$Autonomy = 1$ to $auton$}

  \STATE $SOC \leftarrow SOC - ( 24 - Hours \, of \, PV \, generation)\,h\, discharge$
  \STATE $AutonomyCount \leftarrow AutonomyCount+1)$
  \\  
%    \COMMENT {autonomy verification during discharge period}
%  \\
  \STATE assert NOT($Battery \, empty$)  
  \\
  \COMMENT {Perform similar \textbf{for}-LOOP as defined in line 2-21}
  %%\ENDFOR
  \ENDWHILE
 \RETURN $(\,)$ 
 \end{algorithmic} 
 \label{alg:verification-algorithm}
 \end{algorithm}

\section{Experimental Evaluation} 
\label{sec:results}
%------------------------------------------
\subsection{Description of the Case Studies}
%------------------------------------------
We have performed five case studies to evaluate the tools: (a) four PV systems (three in series 325W PV panels, four 220 Ah batteries in a configuration with two series and two parallel with 48h autonomy, 700 W inverter with surge of 1,600W, charge controller with MPPT with 35A/150V of capacity) deployed in four different houses in an indigenous community (GPS coordinates 2$^{o}$44'50.0"S 60$^{o}$25'47.8"W) situated nearby Manaus (Brazil), with each house having a different power demand (house 1 = 253 W, house 2 = 263 W, house 3 = 283 W, and house 4 = 501 W); and (b) one case concerning a system deployed as an individual system in Manaus (GPS coordinates 3$^{o}$4'20.208"S 60$^{o}$0'30.168"W), supporting 915 W of the house's load (house 5 with four 325W PV panels in a configuration two series and two parallel, four 120Ah batteries in series and autonomy of just 6 h, 1,200 W inverter with surge of 1,600 W, charge controller with MPPT of 150V/35A).

The annual average temperature ($T$) in Manaus is from 23$^{o}$C to 32$^{o}$C; and irradiance ($G$) varies from 274 W/m$^{2}$ to 852 W/m$^{2}$ when there is sunlight (that information is provided in Lines 5 and 6 of Algorithm~\ref{alg:verification-algorithm}). Another characteristic of Manaus, based on historical weather data %\cite{Temperature}, 
\cite{Irradiance}, is related to the fact that only during 8 hours of the day is possible to have PV generation, from 8:00h to 16:00h (that information is provided in Algorithm~\ref{alg:verification-algorithm} as well).

@@

Mostly of experiments were conducted on an otherwise idle Intel octa core Xeon CPU E5-2640 with 2.4 GHz and 264 GB of RAM, running Ubuntu 18.04.1 LTS 64-bits. It was called high-end setup and the verification engines demanded around 90G bytes of RAM (each process).

In addition, it was considered a low-end setup, using specifically the incremental option of ESBMC verification engine, where the demand of memory is very low (2G Bytes of RAM). The experiments were conducted on an otherwise idle Intel Core i7-2600 (8-cores), with 3.4 GHz and 32 GB of RAM, running Ubuntu 18.04.1 LTS 64-bits.

@@

%------------------------------------------
\subsection{Results}
\label{sec:results_indeed}
%------------------------------------------
%
The Table~\ref{tab1} summarize the results concerning the use of the automated verification tools applied to the five case-studies of a solar PV system.

@@

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{makecell}
\usepackage{algorithm,algorithmic}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

@@

\keywords{Automated verification  \and Model Checking \and Electrical systems \and Solar photovoltaic systems.}
\end{abstract}
%
%
%
\section{Introduction}
%%%%%In 2015, 193 Member States of the United Nations agreed as part of the Sustainable Development Goals (SDG) with a specific target to 'ensure access to affordable, reliable and modern energy for all by 2030' – universal access to electricity and clean cooking~\cite{IEAeao2017}. Therefore electricity has a important role, based on two main issues: %that are on the agenda of all current rulers and society in general, regardless of whether it is a developed, emerging or developing country: 
%%%%%the impact on the global climate change and poverty alleviation.
%
%%%%%Related to climate change, the 23rd session of the Conference of the Parties (COP23) to the United Nations Framework Convention on Climate Change (UNFCCC) in 2017 was the second session after the Paris Agreement, which aims to keep the maximum global average temperature rise as close as possible to 1.5 $^{o}$C. %During COP23, with respect to the future market development of renewable power and solar PV (photovoltaics) in particular, it was created an alliance of nations and states committed to moving the world from burning coal to cleaner power sources was a promising sign at COP23 \cite{Jager-Waldau}. 
%%%%%Decarbonization of our energy supply is an important component to achieve the targets, because 65.3\% of the worlds current CO\textsubscript{2} emissions (from energy supple) are due to burning fossil fuels according with 2016 data: 81.1\% of our total primary energy supply depended on burning fossil fuels, namely 27.1\% coal, 31.9\% oil and 22.1\% natural gas~\cite{IEAwes2018}. In terms of final energy consumption electricity only accounted for 18.6\%, but was responsible for 32\% of the total CO\textsubscript{2} emissions~\cite{IEAwes2018}. Photovoltaics (PV) is one of the key technology options for implementing the shift to a decarbonized energy supply, because it is abundant, cannot be monopolized  and can be deployed in a modular way almost everywhere on Earth.
%
%Related to poverty alleviation, 
%
%In reality, the biggest barrier to universal energy access is not the capacity to generate electricity, but rather the ability to get energy to those who need it most. The concept of energy poverty and access tends to focus on very small, incremental shifts in the delivery of energy to poor people. The total energy demand of these shifts constitutes only a small fraction of the ‘gap’ between electricity supply in much of Africa versus more industrialised countries, or the gap between the region’s current and projected electricity supply. https://policy-practice.oxfamamerica.org/static/media/files/FINAL_speakingpowertotruth_SH.pdf
%
%%%%%Poverty, in other hand, needs to be addressed through socio-economic development. Poverty %is conceptualized in material terms as not having access to adequate levels of food, water, clothing, shelter, sanitation, health care and education. This 
%%%%%can be translated into people having insufficient income. 
%
%%%%%%%For numerous people around the world, a better life means getting access to basic needs such as food, health services, housing and clean water. None of these basic needs can be provided without energy. Energy is one of the most essential inputs into sustaining people‘s livelihoods. 
Lack of access to clean and affordable energy is considered a core dimension of poverty~\cite{Hussein2012}. Progress has been made worldwide. The number of people without electricity access fell below 1 billion for the first time in 2017~\cite{IEAweo2018}. To provide universal electricity for all, decentralized systems, led by solar PV in off-grid and mini-grid systems, will be the least-cost solution for three-quarters of the additional connections needed, but that grid extension will still have a role to play, especially in urban areas~\cite{IEAweo2018}.

@@

In each iteration, Automizer pick a trace (sequence of statements) that leads from the initial location to the error location and check whether the trace is feasible or infeasible. If the trace is feasible, an error is reported to the user; otherwise it is computed a sequence of predicates along the trace as a proof of the trace’s infeasibility. The predicates are a sequence of interpolants since each predicate “interpolates” between the set of reachable states and the set of states from which we cannot reach the error. In the refinement step of the CEGAR loop, Automizer tries to find all traces whose infeasibility can be shown with the given predicates and subtract these traces from the set of (potentially spurious) error traces that have not yet been analyzed. It is used automata to represent sets of traces. The difference to a classical CEGAR-based predicate abstraction is that Automizer have to do any logical reasoning (e.g., SMT solver calls) that involves predicates of different CEGAR iterations~\cite{Automizer2018}. 
%
%----------------------------------------
\section{Automated verification of Stand-alone Solar PV Systems }
\label{sec:Methodology}
%----------------------------------------
\subsection{Stand-alone Solar PV Systems}
%%%%%%PV systems are classified into distinct types~\cite{Mohanty}. Specifically for remote rural areas of developing countries or places where the grid extension is not feasible, the most suitable configuration is the regulated stand-alone system with battery and AC load; this configuration is the focus of case-studies of this work.
%%%%%%
The mathematical modeling of the PV system is based on modular blocks, as illustrated in Fig.\ref{fig:blockdiagram}. %It identifies the PV generator, batteries, charge controller, inverter, and AC load. 
The PV generator, which can be a panel or an array, is a semiconductor device that can convert solar energy into DC electricity. In Fig.\ref{fig:blockdiagram}, there are two variables that depend on the site where the system is deployed and the weather (i.e., solar irradiance $G$ and temperature $T$). For night hours or rainy days, we need to hold batteries, where power can be stored and used. The use of battery as a storage form implies the presence of a charge controller~\cite{Hansen}. The PV arrays produce DC and therefore when the PV system contains an AC load, a DC/AC conversion is required. That converter is called of inverter; and the AC load dictates the behavior of AC electrical load from the house that will be fed by the system.
\begin{figure}[h]
\includegraphics[width=0.6\textwidth]{blockdiagramPVS2}
\centering
\caption{Block diagram for a typical stand-alone PV system~\cite{Hansen}.}
\label{fig:blockdiagram} 
\end{figure}
A wide variety of models exists for estimation of power output of PV modules (and $I-V$ or $P-V$ curves). However, the present work will rely on the simplified model of 1-diode. It was demonstrated that it has a small error rate, between 0.03\% and 4.68\% from selected PV panels tested~\cite{Saloux}. The model uses formulas from different references~\cite{Hansen}, \cite{Saloux}, and \cite{Ross}.

To model the batteries, it was used the most common based on lead-acid~\cite{Copetti}. Related to the charge controller, it was used the 4-steps controller processes with MPPT (maximum power point tracking) to decide if the PV array must be disconnected from the system, or if the PV can be reconnected, or if the load is disconnected and when the load is reconnected~\cite{Hansen}. The inverter model is based on efficiency, i.e., the relation between the AC output power and the DC input power of the component~\cite{Hansen}. The AC load, from every case study, is defined by a 24 hour curve who reflects the estimated use of all electrical loads from the house. Load information was collected during a presential survey performed in each house.

@@

Algorithm~\ref{alg:verification-algorithm} describes the pseudo-code used to perform the automated verification. The batteries are assumed to have initial SOC of 80\% (Line 1). Information related to average temperature ($T$) and solar irradiance ($G$), maximum and minimum annual, are given to the algorithm in Lines 7 to 10 using non-deterministic variables from ESBMC to explore all possible states and the \textit{assume} macro to constrain the non-deterministic values using a given range. In order to reduce the computational effort of the algorithm, every 24h-day was considered as a time-step of 1 hour, and it was split into two parts: (a) one where it is possible to occur PV generation, during daylight, but dependent on the sun and weather conditions; and (b) one that includes all the remaining day (without PV generation). Therefore, our approach depends on specific data about the solar irradiation levels to define the average amount of hours of PV generation.

@@

Next, representing the time of the day where PV generation is not possible anymore, starting in Line 22, the algorithm will only discharge the batteries (using the 1 hour step) until a new charging process (at the next day) starts. Specific \textit{assert} in Line 26 will check the state of charging from batteries, and they will violate the property if their levels reach the minimum that represents a discharged battery; therefore, the PV system is unable to supply energy to the house. Nevertheless, if the verification engine does not fail, then we can conclude that the PV system does not need further corrections up to the given bounds.
%
%\textcolor{red}{this sentence is unclear... After this process is started the battery autonomy verification, from line 31}. \textcolor{red}{this sentence is unclear... Based on the fact that won't be PV generation after a given time of the day, the algorithm will only discharge the batteries until a new charging process (at the next day) to start.} \textcolor{red}{what do you mean by The formal verification is guaranteed?...  The formal verification is guaranteed by  macro to specific variables of the model, according lines 27 and 35.}
% and the non-deterministic variables $G$ and $T$ are considered during the formal verification of the system, otherwise, during the other two periods, there is no PV generation and just the power consumption from the backup batteries. 
%Within this 8h-period, $G$ and $T$ are automated verified with different values every one hour.
%, and change their value every 1 h according with the algorithm created using the technique.
 \begin{algorithm}
 \caption{Model checking algorithm for stand-alone PV}
 \begin{algorithmic}[1]
 \renewcommand{\algorithmicrequire}{\textbf{Input:}}
 \renewcommand{\algorithmicensure}{\textbf{Output:}}
  \STATE $SOC \leftarrow 80\%$ \\
%  \COMMENT {Starting with the PV generation time}
% \\ 
 \textit{LOOP Process}
  \FOR {$h = 1$ to $Hours \, of \, PV \, generation$}
  \STATE $G \leftarrow nondet \textunderscore uint(\,)$ \COMMENT {$G$ is non-deterministic variable}
  \STATE $T \leftarrow nondet \textunderscore uint(\,)$ \COMMENT {$T$ is non-deterministic variable}
  \STATE assume ($Gmin \leq G \leq Gmax$) \COMMENT {restricting $G$ values}
  \STATE assume ($Tmin \leq T \leq Tmax$) \COMMENT {restricting $T$ values}
  \STATE $Imax, Vmax \leftarrow PVgenerationMODEL (G,T)$ 
  \\
%  \COMMENT {If-then-else sequence to imitate charge controller work}
  \IF {($battery \, is \, empty$) AND ($PV \, is \, generating$)}
    \STATE charge battery in 1h
    \STATE PV feed the house
  \ELSIF {($battery \, is \, empty$) AND NOT($PV \, is \, generating$)}
    \STATE FAIL with assert macro
  \ELSIF {NOT($battery \, is \, empty$) AND ($PV \, is \, generating$)}
    \STATE stop battery charge
    \STATE PV feed the house
  \ELSIF {NOT($battery \, is \, empty$) AND NOT($PV \, is \, generating$)}
    \STATE discharge battery in 1h
    \STATE Battery feed the house
  \ENDIF
  \STATE $h \leftarrow (h+1)$
  \ENDFOR
 \\ \textit{Start of battery autonomy verification:}
%%  initialization\;
\STATE $AutonomyCount \leftarrow 1$
 \WHILE {$AutonomyCount \leq autonomy$}
  %%\FOR {$Autonomy = 1$ to $auton$}
  \STATE $SOC \leftarrow SOC - ( 24 - Hours \, of \, PV \, generation)\,h\, discharge$
  \STATE $AutonomyCount \leftarrow AutonomyCount+1)$
  \\  
%    \COMMENT {autonomy verification during discharge period}
%  \\
  \STATE assert NOT($Battery \, empty$)  
  \\
  \COMMENT {Perform similar \textbf{for}-LOOP as defined in line 2-21}
  %%\ENDFOR
  \ENDWHILE
 \RETURN $(\,)$ 
 \end{algorithmic} 
 \label{alg:verification-algorithm}
 \end{algorithm}

\textcolor{red}{shall we provide further details about the case studies?}

\section{Experimental Evaluation} 
\label{sec:results}
%------------------------------------------
\subsection{Description of the Case Studies}
%------------------------------------------
We have performed five case studies to evaluate the tools, as described by Table~\ref{tab2}. 
\begin{table}
\caption{Case studies: stand-alone solar PV systems.}\label{tab2}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\hline
Item & House 1 & House 2 & House 3 & House 4 & House 5\\
\hline
\hline
PV Panels &  \multicolumn{4}{|c|}{3x 325W: in series} & 4x 325W: 2x series and 2x parallel \\
\hline
Batteries & \multicolumn{4}{|c|}{\makecell{4x220Ah: 2x series and 2x parallel\\ autonomy: 48h}} & \makecell{4x 120Ah batteries in series\\ autonomy: 6h}\\
\hline
Charge Controller & \multicolumn{5}{|c|}{With MPPT of 150V/35A}\\
\hline
Inverter & \multicolumn{4}{|c|}{700W with surge of 1,600W} & 1,200W with surge of 1,600W\\
\hline
Power demand & 253W & 263W & 283W & 501W & 915W\\
\hline
GPS Coordinates & \multicolumn{4}{|c|}{\makecell{2$^{o}$44'50.0"S 60$^{o}$25'47.8"W\\}} & \makecell{3$^{o}$4'20.208"S \\60$^{o}$0'30.168"W}\\
\hline
Details & \multicolumn{4}{|c|}{\makecell{Riverside indigenous community\\Rural Area of Manaus - Brazil}} & \makecell{Urban house \\Manaus-Amazonas-Brazil}\\
\hline
\hline
\end{tabular}
\end{table}

The annual average temperature ($T$) in Manaus is from 23$^{o}$C to 32$^{o}$C; and irradiance ($G$) varies from 274 W/m$^{2}$ to 852 W/m$^{2}$ when there is sunlight (Lines 5 and 6 of Algorithm~\ref{alg:verification-algorithm}). Another characteristic of Manaus, based on historical weather data %\cite{Temperature}, 
\cite{Irradiance}, is related to the fact that only during 8 hours of the day is possible to have PV generation, from 8:00h to 16:00h. % (that information is provided in Algorithm~\ref{alg:verification-algorithm} as well).

@@

Mostly of experiments were conducted on an otherwise idle Intel octa core Xeon CPU E5-2640 with 2.4 GHz and 264 GB of RAM, running Ubuntu 18.04.1 LTS 64-bits. It was called high-end setup and the verification engines demanded until 90G bytes of RAM.

In addition, it was considered a low-end setup, using specifically the incremental option of ESBMC verification engine, where the demand of memory is low (2G Bytes of RAM). The experiments were conducted on an Intel Core i7-2600 (8-cores), with 3.4 GHz and 32 GB of RAM, running Ubuntu 18.04.1 LTS 64-bits.

@@

\textcolor{red}{What are the command-lines used for all tools? Versions? Solvers? Solvers versions?}

%------------------------------------------
\subsection{Results}
\label{sec:results_indeed}
%------------------------------------------
%
Table~\ref{tab1} summarize the results concerning the use of the automated verification tools applied to the five case-studies of a solar PV system. The time expressed at the Table answer the RQ2. Worth to mention that using simulation tool (HOMER Pro) in all case studies, the 1,200W was the only one that was proved to not meet the requirement of battery autonomy; all the 700W systems had no indication of flaws during simulation. The simulation took less than five seconds to be performed in each case study.

@@

The engines ESBMC and ESBMC Incremental accused that the 1,200W system (House 5) has sizing error related to the right number of solar panels. Related to the 700W systems (Houses 1-4) the verification engines ESBMC and ESBMC incremental indicate that there is a situation of low state of charge from batteries that will cause the cut of the load, just after the 16:00h of the day, when the weather condition was not enough to charge the batteries.

The CBMC tool, was not successful to produce results. It finished the verification of all cases, but the log file indicated that the Solver did not present SAT or UNSAT answer. We tried the MiniSAT and even the Boolector Solvers.

Related to UAutomizer, ....

The flaws from 700W systems were confirmed with the dwellers who own the systems by an interview process and from a local monitoring system with the Raspberry Pi controller: at least once a month is usual the system to turn off, normally in raining or clouds days, further affirming RQ1; related to the 1,200W system, we identified that, in fact, it does not meet the battery autonomy most of the time (mainly when all loads are turned on), and this was double checked with the monitoring system from the charge controller from the manufacturer who showed that the maximum power or surge power were not exceeded, thus affirming RQ1; this behavior is expected since the system was purchased as an off-the-shelf solution and not as a customized design for the electrical charges of the house
